Global search algorithm with local tuning for solving mixed-integer global optimization problems \thanks{This study was supported by the .}

### 2021-04-07 2021/MOTOR_BarkalovLebedev/MOTOR2021.tex
MOTOR правки

@@

\textbf{1. The need to solve multiple global optimization problems.} As mentioned earlier, in the process of solving the MOO problem, it may be necessary to find several different efficient decisions due to possible changes in the optimality requirements. Obtaining different efficient decisions in the proposed approach is ensured by choosing different convolution coefficients (importance indicators) for the objective functions $f_i(y)$, $1 \leq i \leq s$, which results in obtaining different  scalar multi-extremal functions $F(\lambda, y)$ from (\ref{eq:04}). These functions $F(\lambda, y)$ can be optimized sequentially. This determines a multi-stage scheme for solving the MOO problem, when at each stage DM specifies the necessary scalarization coefficients, then the resulting scalar optimization problem is solved, after which DM analyzes the efficient decisions found.

@@

\textbf{2. Step-by-step solution of a set of global optimization problems.} As shown in Section \ref{sec:03}, the information-statistical multiextremal optimization algorithms used in the proposed approach determine the points of consecutive iterations of the global search taking into account the search information
\begin{equation}
\label{eq:22}
A_k=\{(x_i,z_i,f_i )^T:1 \leq i \leq k\}
\end{equation}
obtained in the calculations (see (\ref{eq:14})--(\ref{eq:17})). In (\ref{eq:22}) $x_i$, $1 \leq i \leq k$,  are the reduced points of performed global search iterations ordered in ascending order of coordinates,  $z_i$, $f_i$, $1 \leq i \leq k$, are values of the scalar function $F(\lambda, y)$, from (\ref{eq:04}) and objective functions $f(y)$ from (\ref{eq:01}) of the current optimization problem to be solved at the points $x_i=y(x_i)$, $1 \leq i \leq k$. By using search information $A_k$ from (\ref{eq:22}) when choosing next search iterations it is possible to solve global optimization problems more efficiently and to provide convergence of algorithms only to the global minima of multiextremal functions being minimized.

@@

\textbf{3. Simultaneous solution of a set of global optimization problems.} Information connectivity makes it possible to propose a more general scheme for simultaneous optimization of all the functions of the set $\Phi_p(y)$ from (\ref{eq:21}). In this case, the search information $A_k$ from (\ref{eq:22}) will contain the computed values of all simultaneously optimized functions $F(\lambda_i, y)$, $1 \leq i \leq p$, i.e. 
\begin{equation}
\label{eq:24}
A_k=\{(x_i,\overrightarrow{z_i},f_i )^T : 1 \leq i \leq k\},
\end{equation}
where the values of $\overrightarrow{z_i}$, $1 \leq i \leq k$, represent vectors 
\begin{equation}
\label{eq:25}
\overrightarrow{z_i}=( z_i (1),z_i (2), \dots ,z_i (p)), \; z_i (j)=F(\lambda_j,y(x_i)), 1\leq i \leq k, 1\leq j\leq p.
\end{equation}

@@

The results of numerical experiments are presented in Table \ref{tab:2}. The table shows in the first column the number of functions $\Phi_p (y)$ from (\ref{eq:21}) being optimized.  The columns named Iters contain the average number of iterations performed by the algorithm to solve the MOO problem. The columns named HV and DU contain the values of the HV and DU indicators. The results of the numerical experiments are divided with respect to the three algorithms used: GSA is described in Section \ref{sec:03}, MGSA is discussed in Section \ref{sec:03} (item 2), SGSA is presented in Section \ref{sec:04} (item 3).

@@

@@

@@

>>>>

This paper presents the results of the authors continuing research aimed to develop effective methods for solving MOO problems \cite{c35,c36,c37,c44}. In \cite{c35}, some methods for solving global optimization problems are considered, in \cite{c36},  a general decision-making model based on MOO problems is proposed, and in \cite{c37}, a theoretical analysis is presented regarding the effectiveness of the use of search information when solving MOO problems. In \cite{c44}, we consider MOO problems in which the efficiency criteria can be ordered by importance (lexicographic multi-objective optimization). In this case, MOO problems are reduced to global optimization problems with non-convex constraints for which constrained optimization methods should be applied. In this paper, we propose new computational schemes for using search information that significantly reduce the amount of calculations required to solve MOO problems.

@@

In this Section, the proposed approach for finding $p$, $p>1$ Pareto-optimal decisions of the MOO problem is presented. This problem is formulated as the problem of solving the set $\Phi_p (y)$, $p>1$ of global optimization problems (see \ref{subsec:41}). To solve this set of optimization problems, two efficient computational schemes are proposed. The first scheme is based on the traditional sequential procedure for solving multiple problems, however, to solve each subsequent optimization problem, all the search information obtained during previous calculations is taken into account. This accumulated search information can significantly reduce the number of global search iterations required to solve the next optimization problems of the set $\Phi_p (y)$ (see \ref{subsec:42} and Chapter \ref{sec:05}). In the second scheme, all the problems of the set $\Phi_p (y)$ are solved simultaneously, but the search information obtained when solving each problem is also used for solving all the other problems of the set $\Phi_p (y)$ (see \ref{subsec:43}). This useful exchange of search information also significantly reduces the amount of required calculations for solving the problems of the set $\Phi_p (y)$.

\subsection{The need to solve multiple global optimization problems.} 
\label{subsec:41}

As mentioned earlier, in the process of solving the MOO problem, it may be necessary to find several different efficient decisions due to possible changes in the optimality requirements. Obtaining different efficient decisions in the proposed approach is ensured by choosing different convolution coefficients (importance indicators) for the objective functions $f_i(y)$, $1 \leq i \leq s$, which results in obtaining different  scalar multi-extremal functions $F(\lambda, y)$ from (\ref{eq:04}). These functions $F(\lambda, y)$ can be optimized sequentially. This determines a multi-stage scheme for solving the MOO problem, when at each stage DM specifies the necessary scalarization coefficients, then the resulting scalar optimization problem is solved, after which DM analyzes the efficient decisions found.

@@

\subsection{Step-by-step solution of a set of global optimization problems.} 
\label{subsec:42}

As shown in Section \ref{sec:03}, the information-statistical multiextremal optimization algorithms used in the proposed approach determine the points of consecutive iterations of the global search taking into account the search information
\begin{equation}
\label{eq:22}
A_k=\{(x_i,z_i,f_i )^T:1 \leq i \leq k\}
\end{equation}
obtained in the calculations (see (\ref{eq:14})--(\ref{eq:17})). In (\ref{eq:22}) $x_i$, $1 \leq i \leq k$,  are the reduced points of performed global search iterations ordered in ascending order of coordinates,  $z_i$, $f_i$, $1 \leq i \leq k$, are values of the scalar function $F(\lambda, y)$, from (\ref{eq:04}) and objective functions $f(y)$ from (\ref{eq:01}) of the current optimization problem to be solved at the points $x_i=y(x_i)$, $1 \leq i \leq k$. By using search information $A_k$ from (\ref{eq:22}) when choosing next search iterations it is possible to solve global optimization problems more efficiently and to provide convergence of algorithms only to the global minima of multiextremal functions being minimized.

@@

\subsection{Simultaneous solution of a set of global optimization problems.} 
\label{subsec:43}

Information connectivity makes it possible to propose a more general scheme for simultaneous optimization of all the functions of the set $\Phi_p(y)$ from (\ref{eq:21}). In this case, the search information $A_k$ from (\ref{eq:22}) will contain the computed values of all simultaneously optimized functions $F(\lambda_i, y)$, $1 \leq i \leq p$, i.e. 
\begin{equation}
\label{eq:24}
A_k=\{(x_i,\overrightarrow{z_i},f_i )^T : 1 \leq i \leq k\},
\end{equation}
where the values of $\overrightarrow{z_i}$, $1 \leq i \leq k$, represent vectors 
\begin{equation}
\label{eq:25}
\overrightarrow{z_i}=( z_i (1),z_i (2), \dots ,z_i (p)), \; z_i (j)=F(\lambda_j,y(x_i)), 1\leq i \leq k, 1\leq j\leq p.
\end{equation}

@@

The results of numerical experiments are presented in Table \ref{tab:2} and fig. \ref{fig:1}. The table shows in the first column the number of functions $\Phi_p (y)$ from (\ref{eq:21}) being optimized.  The columns named Iters contain the average number of iterations performed by the algorithm to solve the MOO problem. The columns named HV and DU contain the values of the HV and DU indicators. The results of the numerical experiments are divided with respect to the three algorithms used: GSA is described in Section \ref{sec:03}, MGSA is discussed in Section \ref{sec:03} (item 2), SGSA is presented in Section \ref{sec:04} (item 3).

@@

\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{fig_t2}
  \caption{The average number of iterations for solving bi-criteria two-dimensional MOO problems}
  \label{fig:1}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{fig_t5}
  \caption{The average number of iterations for solving bi-criteria five-dimensional MOO problems}
  \label{fig:1}
\end{figure}

@@

\bibitem{c44} Gergel, V., Kozinov, E., Barkalov, K.: Computationally efficient approach for solving lexicographic multicriteria optimization problems. Optim. Lett. (2020). \doi{10.1007/s11590-020-01668-y}
